Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2015-04-06 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
Continued editting ch 4

\section{Experimental Arrangement}
This series of experiments combined the instrumentation used in the non-combustible wall experiments with gypsum wallboard, which has a combustible paper surface. These experiments enable us to examine our ignition/no ignition predictions from the previous experiments.  A free standing, steel frame with a panel of gypsum wallboard attached to the surface was instrumented to measure the thermal exposure of the wall from the fire in terms of temperature and heat flux.  The fires were initially placed at predetermined, fixed positions away from the wall. The burner was moved toward the wall until the burner was against the wall.  If the results indicated that no burn pattern would result from a given combination of fuel and position, then that fuel and position combination were eliminated from the remainder of the experiments in the study.

@@

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{../Figures/Instrumented_GB_Wall}\\
	\caption[Diagram of the instrumented non-combustible wall]{Diagram of the instrumented non-combustible wall. Two vertical arrays of total heat flux gauges and thermocouples were installed along the centerline of the burner and along the right edge of the burner.}
	\label{Instrumented_GB_Wall}
\end{figure}

@@

In this chapter, as in the previous chapter the averaged peak values for temperatures and heat fluxes will presented at varying distances from the wall. Figures~\ref{IWGB_NG_Temp_FrontSet},\ref{IWGB_NG_Temp_RearCenter} and \ref{IWGB_NG_HF_RearCenter} show the values for the natural gas burner experiments with the gypsum wallboard.

@@

Comparisons NG to Gas to PUF

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/IWGB_NG_TC_Surface_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGB_GAS_TC_Surface_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGB_PUF_TC_Surface_Center_Avg}\\
	\caption[Averaged centerline temperature for the natural gas, gasoline, and foam fires]{Averaged centerline temperature for the natural gas, gasoline, and foam fires. The number following the TC represents the height of thermocouple in decimeters  above the surface of the burner.}
	\label{IWGB_Temp_Comp_FrontCenter}
\end{figure}

@@

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/IWGB_NG_HF_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGB_GAS_HF_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGB_PUF_HF_Center_Avg}\\
	\caption[Averaged centerline temperature for the natural gas, gasoline, and foam fires]{Averaged centerline temperature for the natural gas, gasoline, and foam fires. The number following the TC represents the height of thermocouple in decimeters  above the surface of the burner.}
	\label{IWGB_HF_Comp_Center}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/IWGB_NG_HF_Offset_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGB_GAS_HF_Offset_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGB_PUF_HF_Offset_Avg}\\
	\caption[Averaged centerline temperature for the natural gas, gasoline, and foam fires]{Averaged centerline temperature for the natural gas, gasoline, and foam fires. The number following the TC represents the height of thermocouple in decimeters  above the surface of the burner.}
	\label{IWGB_HF_Comp_Edge}
\end{figure}

@@

@@

The rear wall of the compartment was constructed in the same manner for a given set of experiments. The burner was centered on the middle panel of the rear wall.  As in the fire source characterization experiments the top of the fuel in each case was positioned 7.5~cm above the floor of the compartment.  Each of the painted gypsum board panels that were used for a fire pattern had a 5~cm grid drawn on it to assist with the determination of the geometry of the fire pattern.

>>>>

\section{Experimental Arrangement}
This series of experiments combined the instrumentation used in the non-combustible wall experiments with gypsum wallboard, which has a combustible paper surface. These experiments were conducted in a similar manner as the non-combustible instrumented wall experiments, the only new variable is the painted, gypsum wallboard. These experiments enable us to examine our ignition/no ignition predictions from the previous experiments.  A free standing, steel frame with a panel of gypsum wallboard attached to the surface was instrumented to measure the thermal exposure of the wall from the fire in terms of temperature and heat flux.  The fires were initially placed at predetermined, fixed positions away from the wall. The burner was moved toward the wall until the burner was against the wall.  If the results indicated that no burn pattern would result from a given combination of fuel and position, then that fuel and position combination were eliminated from the remainder of the experiments in the study.

@@

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{../Figures/Instrumented_GB_Wall}\\
	\caption[Diagram of the instrumented gypsum wallboard]{Diagram of the instrumented gypsum wallboard. Two vertical arrays of total heat flux gauges were installed along the centerline of the burner and along the right edge of the burner. A single vertical array of thermocouples was installed centered between the heat flux gauges.}
	\label{Instrumented_GB_Wall}
\end{figure}

@@

The complete set of results from the instrumented gypsum wallboard experiments are provided in Appendix C.  Table ~\ref{tab:IWGB_experiments}    lists the experiments that were conducted for each fuel type.  In this chapter, as in the previous chapter the averaged peak values for temperatures and heat fluxes will presented at varying distances from the wall.  For the gasoline and polyurethane fires, an additional position, 0.08~m from the wall was added.

Figures~\ref{IWGB_NG_Temp_FrontSet},\ref{IWGB_NG_Temp_RearCenter} and \ref{IWGB_NG_HF_RearCenter} show the values for the natural gas burner experiments with the gypsum wallboard.  Revise to match Ch 3 set

\begin{table}
	\centering
	\begin{tabular}{|c|c|c|c|}
		\hline Distance from wall  	& Natural Gas 		& Gasoline			& Polyurethane Foam \\
		\hline (m) 					& Number of Tests 	& Number of Tests  	& Number of Tests 	\\
		\hline 0.61 				& 0 				& 1 				& 2 			 	\\
		\hline 0.31					& 3	 				& 1					& 2 			 	\\
		\hline 0.15					& 1				 	& 1					& 2 			 	\\
		\hline 0.08					& 0 				& 1 				& 2 	 			\\
		\hline 0.00					& 1 				& 1 				& 3 	 			\\
		\hline
	\end{tabular}
	\caption[Matrix of instrumented gypsum wallboard experiments]{Matrix of instrumented gypsum wallboard experiments}
	\label{tab:IWGB_experiments}
\end{table}

@@

The averaged peak temperatures of the wall mounted thermocouples for the three different fuels are shown in Figure~\ref{IWGB_Temp_Comp_FrontCenter}.

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/IWGB_NG_TC_Surface_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGB_GAS_TC_Surface_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGB_PUF_TC_Surface_Center_Avg}\\
	\caption[Averaged centerline temperature for the natural gas, gasoline, and foam fires]{Averaged centerline temperatures, beginning with the top graph position and moving down, the graphs for natural gas, gasoline, and polyurethane foam fires are presented.}  
	\label{IWGB_Temp_Comp_FrontCenter}
\end{figure}

@@

Figures~\ref{IWGB_HF_Comp_Center} and \ref{IWGB_HF_Comp_Edge} present the averaged peak values from the two heat flux gauge arrays.  For the gasoline fire experiment at the 0.08~m position the wall temperatures and the heat flux values from the burner edge array had values that were higher than when those measured at the 0.0 m position. The peak heat release rates for the positions at 0.0~m, 0.08~m and 0.15~m were 120~kW, 110~kW, and 110~kW respectively.  So it does not seem that a change in HRR was the source of the increase.

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/IWGB_NG_HF_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGB_GAS_HF_Center_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGB_PUF_HF_Center_Avg}\\
	\caption[Averaged burner centerline heat flux for the natural gas, gasoline, and foam fires]{Averaged burner centerline heat flux, beginning with the top graph position and moving down, the graphs for natural gas, gasoline, and polyurethane foam fires are presented.}
	\label{IWGB_HF_Comp_Center}
\end{figure}

\begin{figure}[p]
	\centering
	\includegraphics[width=3.5in]{../Figures/IWGB_NG_HF_Offset_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGB_GAS_HF_Offset_Avg}\\
	\includegraphics[width=3.5in]{../Figures/IWGB_PUF_HF_Offset_Avg}\\
	\caption[Averaged burner edge heat flux for the natural gas, gasoline, and foam fires]{Averaged burner edge heat flux for the natural gas, gasoline, and foam fires, from the top down.}
	\label{IWGB_HF_Comp_Edge}
\end{figure}

@@

trim photo sets added here and compare with temps and HF

@@

The burner was centered on the middle panel of the rear wall. The wall behind the center panel of the rear wall was insulated to met the needs of the experiment.  As in the fire source characterization experiments the top of the fuel in each case was positioned 7.5~cm above the floor of the compartment.  Each of the painted gypsum board panels that were used for a fire pattern had a 5~cm grid drawn on it to assist with the determination of the geometry of the fire pattern.
